---
title: "CPFV_pelagics_MPA_block_visualization"
output: github_document
---
project: 2023 cpfv rec fishing and pelagics re: response to letters as part of MPA review
PI: Jenny Selgrath
Purpose of code: to create graphs (bar and line graphs) depicting pelagic CPFV catch in the CDFW fishing blocks associated with Gull Island, Santa Barbara, and Footprint MPAs. These MPAs were specifically identified as being locations that recreational (and commercial) fishermen would like to target pelagic species. 
______________
data input: CPFV_pelagic_species_subset.Rds, MPA_footprint_gullisland_sb_commercial_blocks.shp, CINMS.shp
** see read.me file in Data and Bin folders for more information about data and code. Code adapted/pulled from existing code provided by Chris Free (UCSB) via email on 10/2/2023
_______________
Graph output:  MPA_bar_pelagic.jpeg, MPA_line_pelagic.jpeg, MPA_stacked_gas_all.jpeg, MPA_location_map.jpeg, 
_______________

#Inital setup
## Loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(FishLife)
library(freeR)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(sf)
library(tidyr)
library(viridis)
library(quantmod) 
library(egg)
library(ggspatial)
library(OneR)
library(gganimate)
library(transformr)
library(maps)
library(rnaturalearthdata)
library(rnaturalearth)
path_out="\\\\10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data"
path_plot="\\\\10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/graphs/MPA"
```

## Importing data
```{r cars}
CPFV_pelagics<-readRDS("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/CPFV_pelagic_species_subset.Rds")
MPA_shapefile<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/MPA_fishingblocks/MPA_footprint_gullisland_sb_commercial_blocks.shp")
CINMS<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/CINMS_shapefile/CINMS.shp")
CINMS_surrounding_fishing_blocks<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/MPA_fishingblocks/CINMS_surrounding_fishing_blocks.shp")
gas_adjusted2020<-read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/gas_adjusted2020.csv")
footprint_gull_sb_state_MPA<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/Footprint_gull_SB_state_mpa/Footprint_gull_sb_state_MPA_boundaries.shp")
footprint_gull_sb_fed_MPA<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/Footprint_gull_SB_fed_mpa/Footprint_gull_SB_fed_MPA_boundaries.shp")
footprint_gull_sb_surrounding_blocks<-st_read("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/footprint_gull_SB_surrounding_fishing_blocks/MPA_surrounding_fishing_blocks.shp")
designation_MHW_ENSO_years<-read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/Geom_vlines_designation_MHW_ENSO.csv")
strong_1_pdo<-read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/strong_1_pdo.csv")
```

#______________
#Pre-processing
##Subsetting by MPA blocks
Footprint= 707 and 708
Gull island= 710 and 709
Santa Barbara= 764 and 765
```{r}
CPFV_pelagics_mpa<- CPFV_pelagics%>%
  subset(block_id==707|
           block_id==708|
           block_id==710|
           block_id==709|
           block_id==764|
           block_id==765)
CPFV_pelagics_mpa$year<-as.factor(CPFV_pelagics_mpa$year)
CPFV_pelagics_footprint<-CPFV_pelagics%>% subset(block_id==707|block_id==708)
CPFV_pelagics_gullisland<- CPFV_pelagics%>% subset(block_id==710|block_id==709)
CPFV_pelagics_santabarbara<- CPFV_pelagics%>% subset(block_id==764|block_id==765)
```

##Removing billfish, baitfish, opah, and sharks
```{r}
##removing species
CPFV_pelagics_mpa_rm<-subset(CPFV_pelagics_mpa, CPFV_pelagics_mpa$species_type != "Opah"& CPFV_pelagics_mpa$species_type != "Shark"& CPFV_pelagics_mpa$species_type != "Baitfish"& CPFV_pelagics_mpa$species_type != "Billfish")
```

##Aggregating by year/CPUE
```{r}
mpa_sum_years<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year), FUN=sum)
mpa_sum_years<-rename(mpa_sum_years, "total_catch"= "x")
mpa_count_years<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year), FUN=length)
mpa_count_years<-rename(mpa_count_years, "number_of_trips"= "x")

pelagicsmpa_year_rm<-mpa_sum_years%>% left_join(mpa_count_years, by= "year", relationship= "one-to-one", na_matches = "never")

pelagicsmpa_year_rm<-pelagicsmpa_year_rm %>%
  mutate(CPUE= total_catch/number_of_trips)
```

##Aggregating by year and block/CPUE
```{r}
##aggregating by block and year
mpa_sum_years_block<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year, block_id=CPFV_pelagics_mpa_rm$block_id), FUN=sum)
mpa_sum_years_block<-rename(mpa_sum_years_block, "total_catch"= "x")
mpa_sum_years_block<- replace(mpa_sum_years_block, is.na(mpa_sum_years_block), 0)

mpa_count_years_block<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year, block_id=CPFV_pelagics_mpa_rm$block_id), FUN=length)
mpa_count_years_block<-rename(mpa_count_years_block, "number_of_trips"= "x")
mpa_count_years_block<- replace(mpa_count_years_block, is.na(mpa_count_years_block), 0)


pelagicsmpa_year_block_rm<-mpa_sum_years_block%>% left_join(mpa_count_years_block, by= c("year", "block_id"), relationship = "one-to-one", na_matches = "never")

pelagicsmpa_year_block_rm<-pelagicsmpa_year_block_rm %>%
  mutate(CPUE= total_catch/number_of_trips)
```


##Aggregating by species, year, block/CPUE
```{r}
mpa_sum_year_block_species<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year, species=CPFV_pelagics_mpa_rm$species_name, species_type=CPFV_pelagics_mpa_rm$species_type, block_id= CPFV_pelagics_mpa_rm$block_id), FUN=sum)
mpa_sum_year_block_species<-rename(mpa_sum_year_block_species, "total_catch"= "x")
mpa_sum_year_block_species$year<-as.character(mpa_sum_year_block_species$year)

mpa_count_year_block_species<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(year=CPFV_pelagics_mpa_rm$year, species=CPFV_pelagics_mpa_rm$species_name, species_type=CPFV_pelagics_mpa_rm$species_type, block_id= CPFV_pelagics_mpa_rm$block_id), FUN=length)
mpa_count_year_block_species<-rename(mpa_count_year_block_species, "number_of_trips"= "x")
mpa_count_year_block_species$year<-as.character(mpa_count_year_block_species$year)


pelagicsmpa_species_year_block_rm<-mpa_sum_year_block_species%>% left_join(mpa_count_year_block_species, by= c("year", "block_id", "species", "species_type"), relationship = "one-to-one", na_matches = "never")

pelagicsmpa_species_year_block_rm<-pelagicsmpa_species_year_block_rm %>%
  mutate(CPUE= total_catch/number_of_trips)
```

##Aggregating by species type, year
```{r}
mpa_sum_year_species<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(species_type=CPFV_pelagics_mpa_rm$species_type, year=CPFV_pelagics_mpa_rm$year), FUN=sum)
mpa_sum_year_species<-rename(mpa_sum_year_species, "total_catch"= "x")

mpa_count_year_species<-aggregate(CPFV_pelagics_mpa_rm$total_caught, by=list(species_type=CPFV_pelagics_mpa_rm$species_type, year=CPFV_pelagics_mpa_rm$year), FUN=length)
mpa_count_year_species<-rename(mpa_count_year_species, "number_of_trips"= "x")

pelagicsmpa_species_year_type_rm<-mpa_sum_year_species%>% left_join(mpa_count_year_species, by= c("year", "species_type"), relationship = "one-to-one", na_matches = "never")

pelagicsmpa_species_year_type_rm<-pelagicsmpa_species_year_type_rm%>%
  mutate(CPUE= total_catch/number_of_trips)
```

#_______________
#Graphing
##Bar graph- species type, year
Years with PDO over +1 are 1983, 1986, 1987, 1993, 1997
MHWs with an average intensity over 2 STEVs were in 2015, 2006, and 1995
```{r}
pelagicsmpa_species_year_rm$year<-as.numeric(pelagicsmpa_species_year_rm$year)
ggplot(pelagicsmpa_species_year_rm, aes(year, total, fill=species_type))+ 
  #geom_point()+
  geom_col()+
  #facet_wrap(~species_group)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  theme(legend.position="top")+
  theme(legend.title=element_blank())+
  xlab("Year")+
  ylab("Total Caught")+
  scale_x_continuous(breaks=seq(1980,2020,1))+
  theme(legend.text=element_text(size=7))+
  #adding lines for mpa designation years
  geom_vline(aes(xintercept = 2004), pelagicsmpa_species_year_rm, colour="black", size=1)+
  geom_vline(aes(xintercept = 2007), pelagicsmpa_species_year_rm, colour="black", size=1)+
  geom_vline(aes(xintercept = 2015), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=2)+
  geom_vline(aes(xintercept = 2006), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=2)+
  geom_vline(aes(xintercept = 1995), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=2)+
  #adding lines where ENSO over 1.5 
  annotate("rect", xmin = 1982, xmax = 1983, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .2)+
  annotate("rect", xmin = 1986, xmax = 1987, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .2)+
  annotate("rect", xmin = 1991, xmax = 1993, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .2)+
  annotate("rect", xmin = 1997, xmax = 1998, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .2)+
  annotate("rect", xmin = 2015, xmax = 2016, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .2)+
   #adding shading where PDO is over 1 
  geom_vline(aes(xintercept = 1983), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=3)+
  geom_vline(aes(xintercept = 1986), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=3)+
  geom_vline(aes(xintercept = 1993), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=3)+
  geom_vline(aes(xintercept = 1997), pelagicsmpa_species_year_rm, colour="black", size=0.75, linetype=3)+
  labs(title = str_wrap("CPFV catch overtime in CDFW fishing blocks encompassing Footprint, Gull island, and Santa Barbara MPAs within CINMS", 70))+
   theme(plot.title = element_text(hjust = 0.5))

ggsave("MPA_bar_pelagic.jpeg", width= 30, height= 20, units= "cm", path = path_plot)

```

##Line Plot- species type, year
###PDO color coded -- +1/-1 
```{r}
##Facet wrapped 
pelagicsmpa_species_year_type_rm$year<-as.character(pelagicsmpa_species_year_type_rm$year)
pelagicsmpa_species_year_type_rm$year<-as.numeric(pelagicsmpa_species_year_type_rm$year)
designation_MHW_ENSO_years$year<-as.numeric(designation_MHW_ENSO_years$year)

ggplot()+ 
  geom_line(data=pelagicsmpa_species_year_type_rm, aes(x=year, y=CPUE),size=1, color= "black")+
  facet_wrap(~species_type)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  theme(legend.position="top")+
  theme(legend.title=element_blank())+
  ylab("Catch per Unit Effort (fish caught per trip)")+
  xlab("Years")+
  scale_x_continuous(breaks=seq(1980,2020,5))+
  theme(legend.text=element_text(size=7))+
  theme(strip.text.x = element_text(size = 15))+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="MHW"|Event== "ENSO"), 
           mapping= aes(xintercept=year), color="gray61",  linetype="dotted", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="Federal_MPA_Designation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="State_MPA_Desingation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
   annotate("rect", xmin = 1983, xmax = 1984, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1986, xmax = 1988, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1993, xmax = 1994, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1997, xmax = 1998, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1999, xmax = 2002, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
   annotate("rect", xmin = 2008, xmax = 2014, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)

ggsave("MPA_line_pelagic_strong_pdo.jpeg", width= 30, height= 20, units= "cm", path = path_plot)
```
###PDO color coded-- All non-neutral PDO
```{r}
##Facet wrapped 
pelagicsmpa_species_year_type_rm$year<-as.character(pelagicsmpa_species_year_type_rm$year)
pelagicsmpa_species_year_type_rm$year<-as.numeric(pelagicsmpa_species_year_type_rm$year)
designation_MHW_ENSO_years$year<-as.numeric(designation_MHW_ENSO_years$year)

ggplot()+ 
  geom_line(data=pelagicsmpa_species_year_type_rm, aes(x=year, y=CPUE),size=1, color= "black")+
  facet_wrap(~species_type)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  theme(legend.position="top")+
  theme(legend.title=element_blank())+
  ylab("Catch per Unit Effort (fish caught per trip)")+
  xlab("Years")+
  scale_x_continuous(breaks=seq(1980,2020,5))+
  theme(legend.text=element_text(size=7))+
  theme(strip.text.x = element_text(size = 15))+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="MHW"|Event== "ENSO"), 
           mapping= aes(xintercept=year), color="gray61",  linetype="dotted", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="Federal_MPA_Designation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="State_MPA_Desingation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
   annotate("rect", xmin = 1980, xmax = 1982, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1983, xmax = 1985, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1986, xmax = 1988, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1992, xmax = 1994, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1995, xmax = 1998, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 2003, xmax = 2004, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
  annotate("rect", xmin = 2014, xmax = 2017, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
  annotate("rect", xmin = 1982, xmax = 1983, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1989, xmax = 1992, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1994, xmax = 1995, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1998, xmax = 2003, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 2004, xmax = 2014, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 2018, xmax = 2020, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)
ggsave("MPA_line_pelagic_all_pdo.jpeg", width= 30, height= 20, units= "cm", path = path_plot)
```


##Line Plot- All catch, year, gas, map
Couldn't get the dimesions to graph correctly on the 3 level stacked graph, so making the location graph seperately and then joining outside of R 
###Stacked Line Graph and gas prices 
```{r}
pelagicsmpa_year_rm$year<-as.numeric(as.character(pelagicsmpa_year_rm$year))
plot1<- ggplot()+ 
  geom_line(data=pelagicsmpa_year_rm, aes(x=year, y=CPUE), size=1, color="black")+
  #facet_wrap(~species_type)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  theme(legend.position="top")+
  theme(legend.title=element_blank())+
  ylab("Catch per Unit Effort (fish caught per trip)")+
  xlab("Years")+
  scale_x_continuous(breaks=seq(1980,2020,5))+
  theme(legend.text=element_text(size=7))+
  theme(strip.text.x = element_text(size = 15))+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="MHW"|Event== "ENSO"), 
           mapping= aes(xintercept=year), color="gray61",  linetype="dotted", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="Federal_MPA_Designation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
  geom_vline(data=subset(designation_MHW_ENSO_years, Event=="State_MPA_Desingation"), 
             mapping= aes(xintercept=year),color="gray41", size=1)+
   annotate("rect", xmin = 1980, xmax = 1982, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1983, xmax = 1985, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1986, xmax = 1988, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1992, xmax = 1994, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 1995, xmax = 1998, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
   annotate("rect", xmin = 2003, xmax = 2004, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
  annotate("rect", xmin = 2014, xmax = 2017, ymin = -Inf, ymax = Inf, fill= "red",
        alpha = .1)+
  annotate("rect", xmin = 1982, xmax = 1983, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1989, xmax = 1992, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1994, xmax = 1995, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 1998, xmax = 2003, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 2004, xmax = 2014, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)+
  annotate("rect", xmin = 2018, xmax = 2020, ymin = -Inf, ymax = Inf, fill= "blue",
        alpha = .1)
plot1

plot2<-ggplot(gas_adjusted2020, aes(Year, adjust_price))+
  geom_line(size=1)+
  theme_bw()+
  theme(plot.margin = unit(c(5,5,5,5), 'points'))+
  scale_x_continuous(breaks=seq(1980,2020,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
   scale_x_continuous(breaks=seq(1980,2020,10))+
  theme(legend.text=element_text(size=7))+
  theme(strip.text.x = element_text(size = 15))+
  labs(y = paste0("Inflation Adjusted Gas Price ($/gallon)"))+
  xlab("Year")
plot2
plotstacked<-ggarrange(plot1, plot2, nrow=2)
plotstacked
ggsave("MPA_stacked_gas_all.jpeg", plot= plotstacked, width = 7, height = 8, units="in", path = path_plot)
```

###Location Map
```{r}
plot3<-ggplot()+ 
  geom_sf(data=CINMS_surrounding_fishing_blocks, fill="lightblue" )+
  geom_sf(data=CINMS, fill=NA)+
  geom_sf(data= MPA_shapefile, fill="darkblue", alpha=0.25, color="black")+
  theme_classic()+
  annotation_scale(location= "bl", pad_y = unit(1, "cm"), pad_x = unit(1, "cm"))+
  annotation_north_arrow(location= "tr", which_north= "true", width= unit(1.5, "cm"), style= north_arrow_fancy_orienteering(), pad_x = unit(1, "cm"))+
  xlab("Longitude")+
  ylab("Latitude")
plot3
ggsave("MPA_location_map.jpeg", plot= plot3, width = 10, units="in", path = path_plot)
```

#_______________
#Heat Map 
##Joining shapefile + data
```{r}
##putting in Zeros where there isn't data about fishing blocks
complete_pelagicsmpa_year_block_rm<-tidyr::complete(pelagicsmpa_year_block_rm, block_id, year, fill=list(total=0))

##joining shapefile and data
MPA_shapefile<-rename(MPA_shapefile, "block_id"= "BLOCK10_ID")
MPA_fishing_blocks_pelagic<-left_join(MPA_shapefile, complete_pelagicsmpa_year_block_rm, by= "block_id", relationship= "many-to-many", na_matches= "never")

##removing NAs
MPA_fishing_blocks_pelagic<- subset(MPA_fishing_blocks_pelagic, !is.na(MPA_fishing_blocks_pelagic$total))
```

###Path
```{r}
path_plot_heatmap<-"\\\\10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/graphs/MPA/MPA_heatmaps_by_year"
```


##"For loop" graphing
iterative loop for printing each block 
###Independent scale
####For before MPA designation state= 2004
```{r}
##Subsetting so can add MPAs as designated
MPA_fishing_blocks_pelagic$year<-as.character(as.factor(MPA_fishing_blocks_pelagic$year))
MPA_fishing_blocks_pelagic_2004<- subset(MPA_fishing_blocks_pelagic, year<2004)


##Graphing and saving
for(i in unique(MPA_fishing_blocks_pelagic_2004$year)){
  x<-ggplot()+
    geom_sf(data=footprint_gull_sb_surrounding_blocks, fill="white")+
    geom_sf(data= subset(MPA_fishing_blocks_pelagic_2004, year==i & year<2004),mapping =aes(fill=CPUE))+
    scale_fill_gradientn(colours=rev(magma(6)),
                         name="Catch per unit effort \n(fish caught/trip)",
                         na.value = "grey100", limits= c(0, 200))+
    theme_classic()+
    annotation_scale(location= "bl", width_hint= 0.2, pad_y = unit(1, "cm"), pad_x = unit(0.7, "cm"))+
    annotation_north_arrow(location= "tr", which_north= "true", width= unit(1, "cm"),  
                           height= unit(1, "cm"),style= north_arrow_fancy_orienteering(), 
                           pad_x = unit(0.5, "cm"), pad_y = unit(0.1, "cm"))+
    xlab("Longitude")+
    ylab("Latitude")+
    ggtitle("CPFV CPUE in Footprint, Gull Island, and Santa Barbara MPAs", 
            subtitle= paste("Year", i, sep = "-"))
  
  ggsave(x, path= path_plot_heatmap, file=paste0("MPA_heatmap_individual_", i, ".png") )
}
```


####For after state designation MPAs
```{r}
##Subsetting so can add MPAs as designated
MPA_fishing_blocks_pelagic_2004_2007<- subset(MPA_fishing_blocks_pelagic, year>2003 & year<2007)

##Graphing and printing
for(i in unique(MPA_fishing_blocks_pelagic_2004_2007$year)){
  x<- ggplot()+
    geom_sf(data=footprint_gull_sb_surrounding_blocks, fill="white")+
    geom_sf(data= subset(MPA_fishing_blocks_pelagic_2004_2007, 
            year==i & year >2003 & year< 2007), aes(fill=CPUE))+
    geom_sf(data=footprint_gull_sb_state_MPA, color="black", fill= "darkblue", alpha=0.25)+
    scale_fill_gradientn(colours=rev(magma(6)),
                         name="Catch per unit effort \n(fish caught/trip)",
                         na.value = "grey100", limits= c(0, 200))+
    theme_classic()+
    annotation_scale(location= "bl", width_hint= 0.2, pad_y = unit(1, "cm"), pad_x = unit(0.7, "cm"))+
    annotation_north_arrow(location= "tr", which_north= "true", width= unit(1, "cm"),  
                           height= unit(1, "cm"),style= north_arrow_fancy_orienteering(), 
                           pad_x = unit(0.5, "cm"), pad_y = unit(0.1, "cm"))+
    xlab("Longitude")+
    ylab("Latitude")+
    ggtitle("CPFV CPUE in Footprint, Gull Island, and Santa Barbara MPAs", 
            subtitle= paste("Year", i, sep = "-"))
  
  ggsave(x, path= path_plot_heatmap, file=paste0("MPA_heatmap_individual_", i, ".png") )
}
```

#### For after fed designation of MPAs
```{r}
##Subsetting so can add MPAs as designated
MPA_fishing_blocks_pelagic_2007<- subset(MPA_fishing_blocks_pelagic, year>2006)

##Graphing and printing
for(i in unique(MPA_fishing_blocks_pelagic_2007$year)){
  x<- ggplot()+
      geom_sf(data=footprint_gull_sb_surrounding_blocks, fill="white")+
      geom_sf(data= subset(MPA_fishing_blocks_pelagic_2007, year==i & year>2006), aes(fill=CPUE))+
      geom_sf(data=footprint_gull_sb_fed_MPA, color="black", fill="darkgreen", alpha= 0.25)+
      geom_sf(data=footprint_gull_sb_state_MPA, color="black", fill= "darkblue", alpha=0.25)+
      scale_fill_gradientn(colours=rev(magma(6)),
                         name="Catch per unit effort \n(fish caught/trip)",
                         na.value = "grey100", limits= c(0, 200))+
    theme_classic()+
    annotation_scale(location= "bl", width_hint= 0.2, pad_y = unit(1, "cm"), pad_x = unit(0.7, "cm"))+
    annotation_north_arrow(location= "tr", which_north= "true", width= unit(1, "cm"),  
                           height= unit(1, "cm"),style= north_arrow_fancy_orienteering(), 
                           pad_x = unit(0.5, "cm"), pad_y = unit(0.1, "cm"))+
    xlab("Longitude")+
    ylab("Latitude")+
    ggtitle("CPFV CPUE in Footprint, Gull Island, and Santa Barbara MPAs", 
            subtitle= paste("Year", i, sep = "-"))
  
  ggsave(x, path= path_plot_heatmap, file=paste0("MPA_heatmap_individual_", i, ".png") )
}
```
