---
title: "CPFV_GAM_SoCal"
author: "Van Deusen"
date: "2026-01-21"
output: html_document
---
project: 2023 cpfv rec fishing and pelagics re: response to letters as part of MPA review
PI: Jenny Selgrath
Purpose of code: to create a generalized additive model (GAM) to determine how the number of fish caught per fisher (CPUE) is effected by temperature, gas prices, year, and MPA designation 

~~It seems to me like it is easiest to fit a GAM for each species, so that species doesn't have to be a consideration within the GAM itself

*Response variable*-- number of fish caught per fisher 
*Predictor variables*-- 
Structural: Year (continuous, circular)
Environmental: Sea Surface Temperature (continuous); Marine Heat Wave Index (continuous); Pacific Decadal Oscillation (continuous)
Economic: Gas prices (continuous)
______________
data input: 
_______________
Graph output:  
_______________
## Loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(freeR)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(sf)
library(tidyr)
library(quantmod) 
library(OneR)
library(mgcv)
library(nlme)
path_out="C:/Usersvanessa.van.deusen/Desktop/Van.Deusen_Local/CPFV_local/Data/Model_outputs"

```

##Importing data 
```{r}
fishperfisher <- readRDS()



```




