---
title: "CPFV_pelagics_data_clean"
author: "Vanessa Van Deusen"
date: "2023-11-06"
output: html_document
---
project: 2023 cpfv rec fishing and pelagics re: response to letters as part of MPA review
PI: Jenny Selgrath
data input: cpfv_merged_no_pii.csv; blocks.csv; MFDE_port_historic.csv; MFDE_Species_historic.csv
** see read.me file in Data and Bin folders for more information about data and code. Code adapted/pulled from existing code provided by Chris Free (UCSB) via email on 10/2/2023
data output: CDFW_1980_2020_cpfv_logbook_data.Rds

#Inital setup
## Loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(FishLife)
library(freeR)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
path_out="\\\\10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data"
```

## Importing data
```{r}
cpfv_merged_no_pii<- read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/cpfv_merged_no_pii.csv")
blocks <- read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/blocks.csv")
port_DF <- read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/MFDE_port_historic.csv")
species_DF<-read_csv("//10.74.13.28/ci-users/Vanessa.Van.Deusen/R_projects/CPFV_pelagics/Data/MFDE_Species_historic.csv")
```

## Formatting data
```{r}
cpfv_merged_no_pii$SpeciesCode<-as.numeric(cpfv_merged_no_pii$SpeciesCode)

block_key <- blocks %>% 
  sf::st_drop_geometry()%>% 
  rename("block_id"= "BLOCK10_ID", "block_areaMi"= "AreaMi", "block_Acres"="Acres", "block_shape_area"= "Shape__Area", "block_shape_length"= "Shape__Length")

port_key<- port_DF %>% 
  clean_names("snake")%>%
  rename(port= port_name, port_complex= port_area)%>%
  select(-c(x5))


species_key<-species_DF %>% 
  clean_names("snake") %>%
  rename("species_code"= "species_id", "species_discontinue_date"= "discontinue_date")%>% 
  select(-c(x6))
species_key$species_code<-as.numeric(species_key$species_code)

```


#-----
#Demographics on species, methods, and bait (dataset from 1980-2020)
## Which species groups are most frequently targeted by recreational fishers?
```{r}
target_n <- rev(sort(nrow(cpfv_merged_no_pii)-freeR::complete(cpfv_merged_no_pii %>% select(TargetSpeciesLingcod:TargetSpeciesMiscBay))))
head(target_n)
```

## Which fishing method is most frequently used by recreational fishers?
```{r}
method_n <- rev(sort(nrow(cpfv_merged_no_pii)-freeR::complete(cpfv_merged_no_pii %>% select(FishingMethodTrolling:FishingMethodOther))))
head(method_n)
```
## Which bait is most frequently used by recreational fishers?
```{r}
bait_n <- rev(sort(nrow(cpfv_merged_no_pii)-freeR::complete(cpfv_merged_no_pii %>% select(BaitUsedAnchoviesLive:BaitUsedOtherDead))))
head(bait_n)
```

##Table format
###Specific species 
```{r}
table(cpfv_merged_no_pii$TargetSpeciesLingcod)
table(cpfv_merged_no_pii$TargetSpeciesOther)
table(cpfv_merged_no_pii$TargetSpeciesRockfishes)
table(cpfv_merged_no_pii$TargetSpeciesSalmon)
table(cpfv_merged_no_pii$TargetSpeciesSharks)
table(cpfv_merged_no_pii$TargetSpeciesStripedBass)
table(cpfv_merged_no_pii$TargetSpeciesSturgeon)
table(cpfv_merged_no_pii$TargetSpeciesTuna)
table(cpfv_merged_no_pii$TargetSpeciesPotluck)
table(cpfv_merged_no_pii$TargetSpeciesMiscCoastal)
table(cpfv_merged_no_pii$TargetSpeciesMiscBay)
table(cpfv_merged_no_pii$TargetSpeciesMiscOffshore)
```

###Specific method
```{r}
table(cpfv_merged_no_pii$FishingMethodTrolling)
table(cpfv_merged_no_pii$FishingMethodAnchored)
table(cpfv_merged_no_pii$FishingMethodDiving)
table(cpfv_merged_no_pii$FishingMethodDrifting)
table(cpfv_merged_no_pii$FishingMethodLightTackle)
table(cpfv_merged_no_pii$FishingMethodMooching)
table(cpfv_merged_no_pii$FishingMethodOther)
```

###Specific bait
```{r}
table(cpfv_merged_no_pii$BaitUsedAnchoviesDead)
table(cpfv_merged_no_pii$BaitUsedAnchoviesLive)
table(cpfv_merged_no_pii$BaitUsedOtherDead)
table(cpfv_merged_no_pii$BaitUsedOtherLive)
table(cpfv_merged_no_pii$BaitUsedSardinesDead)
table(cpfv_merged_no_pii$BaitUsedSardinesLive)
table(cpfv_merged_no_pii$BaitUsedSquidDead)
table(cpfv_merged_no_pii$BaitUsedSquidLive)
```




#-----
#Clean data 
```{r}
 data_clean <- cpfv_merged_no_pii %>% 
  janitor::clean_names("snake") %>% 
  rename(date=log_date,
         month=log_month,         
         day=log_day,
         year=log_year,
         block_id=block,
         hm_fished=hours_minutes_fished,
         comm_name_orig=species, 
         temp_f=temperature,
         depth_ft=depth,
         n_fishers=number_of_fishers,
         hrs_fished=hours_fished,
         n_kept=number_kept,
         n_released=number_released,
         n_lost_to_sea_lions=number_lost_to_sea_lions,
         n_caught_by_crew=number_of_fish_caught_by_crew,
         n_crew_fished=number_of_crew_fished) %>% 
   mutate(date=lubridate::mdy(date),
         date_received=lubridate::mdy(date_received),
         date_submitted=lubridate::mdy(date_submitted)) %>% 
  mutate(across(.cols=month:year, .fns=as.numeric))  %>%
  mutate(port_code=as.numeric(port_code)) %>%
  left_join(port_key, by="port_code", relationship = "many-to-one", na_matches = "never" ) %>%
  mutate(port=ifelse(is.na(port), "Invalid", port)) %>%
  mutate(trip_type=stringr::str_to_sentence(trip_type)) %>%
  mutate(block_id=as.numeric(block_id)) %>%
  left_join(block_key %>% select("block_id", "block_areaMi", "block_Acres"), by="block_id") %>%
  mutate(depth_ft=as.numeric(depth_ft) %>% abs(.)) %>% 
  mutate(temp_f=as.numeric(temp_f))%>%
  mutate(n_fishers=as.numeric(n_fishers), hrs_fished=as.numeric(hrs_fished)) %>%
  #mutate(hm_fished=stringr::str_squish(hm_fished), hm_fished=recode(hm_fished, "800"="", "100"="", "145"="", "600"="", "300"="", "830"="", "730"="", "330"="", "251"="", "615"=""),hm_fished=ifelse(hm_fished=="", NA, hm_fished)) 
  ##I am not sure what this code is doing
  mutate(comm_name_orig=stringr::str_squish(comm_name_orig))%>%
 # mutate(species_code=recode(species_code, "Oarfish"="") %>% as.numeric(.)) %>%
  left_join(species_key %>% select(species_code, species_name, species_group,species_discontinue_date), by=c("species_code")) %>%
  #mutate(species_name=ifelse(species_code==2665 , "Oarfish", species_name))%>%
  mutate(across(.cols=n_kept:n_crew_fished, .fns=as.numeric)) %>%
  mutate(across(.cols=n_kept:n_caught_by_crew, .fns=function(x){ifelse(is.na(x), 0, x)}))%>% 
  mutate(target_species_rockfishes=ifelse(!is.na(target_species_rockfishes), "Rockfish", "NA"),
         target_species_misc_coastal=ifelse(!is.na(target_species_misc_coastal), "Misc. coastal", "NA"),
         target_species_other=ifelse(!is.na(target_species_other), "Other", "NA"),
         target_species_lingcod=ifelse(!is.na(target_species_lingcod), "Lingcod", "NA"),
         target_species_tuna=ifelse(!is.na(target_species_tuna), "Tuna", "NA"),
         target_species_misc_offshore=ifelse(!is.na(target_species_misc_offshore), "Misc. offshore", "NA"),
         target_species_salmon=ifelse(!is.na(target_species_salmon), "Salmon", "NA"),
         target_species_potluck=ifelse(!is.na(target_species_potluck), "Potluck", "NA"),
         target_species_striped_bass=ifelse(!is.na(target_species_striped_bass), "Striped bass", "NA"),
         target_species_sharks=ifelse(!is.na(target_species_sharks), "Sharks", "NA"),
         target_species_sturgeon=ifelse(!is.na(target_species_sturgeon), "Sturgeon", "NA"),
         target_species_misc_bay=ifelse(!is.na(target_species_misc_bay), "Misc. bay", "NA"))%>%
   mutate(target_species=paste(target_species_rockfishes, target_species_misc_coastal, 
                              target_species_other, target_species_lingcod, 
                              target_species_tuna, target_species_misc_offshore, 
                              target_species_salmon, target_species_potluck, 
                              target_species_striped_bass, target_species_sharks, 
                              target_species_sturgeon, target_species_misc_bay, sep=", ")) %>% 
  mutate(target_species=gsub("NA, |, NA", "", target_species)) %>% ## not completely sure how this works ("NA, |, NA")
  select(-c(target_species_lingcod:target_species_misc_bay)) %>%
  mutate(fishing_method_anchored=ifelse(!is.na(fishing_method_anchored), "Anchored", "NA"),
         fishing_method_drifting=ifelse(!is.na(fishing_method_drifting), "Drifting", "NA"),
         fishing_method_trolling=ifelse(!is.na(fishing_method_trolling), "Trolling", "NA"),
         fishing_method_light_tackle=ifelse(!is.na(fishing_method_light_tackle), "Light tackle", "NA"),
         fishing_method_diving=ifelse(!is.na(fishing_method_diving), "Diving", "NA"),
         fishing_method_other=ifelse(!is.na(fishing_method_other), "Other", "NA"),
         fishing_method_mooching=ifelse(!is.na(fishing_method_mooching), "Mooching", "NA")) %>% 
  mutate(fishing_method=paste(fishing_method_anchored, fishing_method_drifting, 
                              fishing_method_trolling, fishing_method_light_tackle, 
                              fishing_method_diving, fishing_method_other, 
                              fishing_method_mooching, sep=", "))%>%
  mutate(fishing_method=gsub("NA, |, NA", "", fishing_method)) %>% 
  select(-c(fishing_method_trolling:fishing_method_other))%>% 
  mutate(bait_used_squid_dead=ifelse(!is.na(bait_used_squid_dead), "Squid (dead)", "NA"),
         bait_used_sardines_live=ifelse(!is.na(bait_used_sardines_live), "Sardines (live)", "NA"),
         bait_used_anchovies_live=ifelse(!is.na(bait_used_anchovies_live), "Anchovies (live)", "NA"),
         bait_used_squid_live=ifelse(!is.na(bait_used_squid_live), "Squid (live)", "NA"),
         bait_used_other_dead=ifelse(!is.na(bait_used_other_dead), "Other (dead)", "NA"),
         bait_used_anchovies_dead=ifelse(!is.na(bait_used_anchovies_dead), "Anchovies (dead)", "NA"),
         bait_used_sardines_dead=ifelse(!is.na(bait_used_sardines_dead), "Sardines (dead)", "NA"), 
         bait_used_other_live=ifelse(!is.na(bait_used_other_live), "Other (live)", "NA")) %>% 
  mutate(bait_used=paste(bait_used_squid_dead, bait_used_sardines_live, 
                         bait_used_anchovies_live, bait_used_squid_live, 
                         bait_used_other_dead, bait_used_anchovies_dead, 
                         bait_used_sardines_dead, bait_used_other_live, sep=", ")) %>% 
  mutate(bait_used=gsub("NA, |, NA", "", bait_used)) %>% 
  select(-c(bait_used_anchovies_live:bait_used_other_dead))%>%
  mutate(ID=seq.int(nrow(data_clean)))%>%
  mutate(logbook_id_use=paste(ID, date, sep="-"))%>%
  select(logbook_id_use, 
         year, month, day, 
         date, date_submitted, date_received,
         port_complex, port_code, port,
         no_activity_month,
         trip_type, non_paying, descending_device, bird_interaction,
         target_species, fishing_method, bait_used, 
         block_id,
         depth_ft, temp_f, 
         departure_time, return_time, hm_fished, hrs_fished, n_fishers, n_crew_fished,
         species_code, species_name,
         n_kept, n_released, n_lost_to_sea_lions, n_caught_by_crew,
         everything())

```

# Build Logbook ID
```{r}
# Build logbook id
log_key <- data_clean %>% 
  group_by(year, date,
           port_complex, port, port_code, 
           no_activity_month, trip_type, non_paying, 
           departure_time, return_time,
           block_id,
           depth_ft, temp_f,
           hm_fished, hrs_fished,
           n_fishers, n_crew_fished, logbook_id_use) %>% 
  summarize(n_kept=sum(n_kept)) %>% 
  ungroup()
# Are the logbook ids unique? Mine should be, there's shouldn't be
freeR::which_duplicated(log_key$logbook_id_use)

# Does number of fishers always exceed number of crew?
sum(log_key$n_crew_fished < log_key$n_fishers, na.rm=T)
```

#Inspect Data
```{r}
# Inspect data
################################################################################

# Inspect
head(data_clean)
str(data_clean)
# na_check <- freeR::complete(data)
# (100 - na_check / nrow(data) * 100) %>% round(., 2)

# Date
range(data_clean$date)
range(data_clean$year)
range(data_clean$month)
range(data_clean$day)
range(data_clean$date_received, na.rm=T)
range(data_clean$date_submitted, na.rm=T) # 1900-01-01 = NA?
range(data_clean$temp_f, na.rm=T)

# Ports
port_key_check <- data_clean %>% 
  select(port_complex, port_code, port) %>% 
  unique() %>% 
  arrange(port_code)

port_invalid<-data_clean%>%
  subset(port=="Invalid")%>%
  select(port_complex, port_code, port)%>%
  unique()%>%
  arrange(port_code)
#write.csv(port_invalid_full, file.path(path_out, "port_invalid_full.csv"))

# Block
block_key_check <- data_clean %>% 
  select(block_id) %>% 
  unique()
sort(block_key_check$block_id[is.na(block_key_check$block_type)])
blocks_wrong<-data_clean%>%
  subset(block_id>897)%>%
  select(block_id, year)%>%
  unique()%>%
  arrange(block_id)
#write.csv(blocks_wrong, file.path(path_out, "blocks_wrong.csv"))

##Looking specifically at one repeatedly misentered block ID that was re-entered 172 times 
block1032_wrong<-data_clean%>% subset(block_id==1032)
write.csv(block1032_wrong, file.path(path_out, "block1032_wrong.csv"))

# Trip type
sort(unique(data_clean$trip_type))
table(data_clean$non_paying) # N, Y
table(data_clean$no_activity_month)

# Descending device
table(data_clean$descending_device) # B, N, Y
table(data_clean$bird_interaction) # B, N, Y

# Species
sort(unique(data_clean$species_code))
sort(unique(data_clean$species_name))
species_key_check <- data_clean %>% 
  select(species_code,species_name, comm_name_orig) %>% 
  unique()
sort(species_key_check$species_code[is.na(species_key_check$species_name)])

species_name_missing<-data_clean%>%
  subset(is.na(data_clean$species_name))%>%
  select(species_code, species_name, comm_name_orig) %>%
 # unique()%>%
  arrange(species_code)
###comparing species names to comm_name_orig
species_key_check_against<-data_clean%>%
  select(species)

###break down of missing species names 
table(species_name_missing$comm_name_orig)
#write.csv(species_name_missing, file.path(path_out, "species_name_missing.csv"))

# Temperature
ggplot(data_clean, aes(y=temp_f)) +
  geom_boxplot() +
  labs(y="Temperature (°F)") +
  lims(y=c(0,100)) +
  theme_bw()

hist(data_clean$temp_f)
temp_mistake<-subset(data_clean, temp_f> 100)
view(temp_mistake)
#write.csv(temp_mistake, file.path(path_out, "temp_mistake.csv"))


# Depth
ggplot(data_clean, aes(y=depth_ft)) +
  geom_boxplot() +
  labs(y="Depth (feet)") +
  lims(y=c(0, 500)) +
  theme_bw()
hist(data_clean$depth_ft)

# Times
freeR::uniq(data_clean$departure_time)
freeR::uniq(data_clean$return_time)
freeR::uniq(data_clean$hm_fished)

# Inspect duration
duration_key <- data_clean %>% 
  select(hm_fished, hrs_fished) %>% 
  unique() %>% 
  arrange(hrs_fished)
#Dont understand what is purpose of below
#duration_key_check <- duration_key %>% 
 # mutate(nchar=nchar(hm_fished),
  #       colon_yn=grepl(":", hm_fished)) %>% 
  #rowwise() %>% 
  #mutate(hrs_fished2=conv_hm_to_hr(hm_fished)) %>% 
#  ungroup() %>% 
 # mutate(hrs_check=near(hrs_fished, hrs_fished2, tol=0.0001))

# Target species, fishing method, bait used
table(data_clean$target_species)
table(data$fishing_method)
table(data$bait_used)

```

# Didn't do-- Free Final processing
```{r}

################################################################################

# Final format
data_out <- data %>% 
  # Add HR calculated from HM
  # (proven to be more comprehensive and identical)
  left_join(duration_key_check %>% select(hm_fished, hrs_fished2), by="hm_fished") %>% 
  select(-hrs_fished) %>% 
  rename(hrs_fished=hrs_fished2) %>% 
  select(filename:hm_fished, hrs_fished, everything()) 

```

#Export
```{r}
saveRDS(data_clean, file.path(path_out, "CDFW_1980_2020_cpfv_logbook_data.Rds"))
```

